Galaxy Clustering Constraints on the Galaxy Size--Halo Connection

### 2017-08-31 paper/galsize_paper.tex
Finished draft of 4.1.2

In this section we present new measurements of the $\rhalf-$dependence of projected galaxy clustering, $\wproj.$ Galaxy clustering has complex, simultaneous dependence upon $\mstar,$ $\ssfr,$ $\rhalf,$ and $\bt.$ Since the purpose of this work is to test models of composite galaxy size $\rhalf,$ we design a measurement that is specifically tailored to isolate the influence of $\rhalf$ on the two-point function, while minimizing the dependence upon the other correlated variables.

To achieve this, we compute projected clustering separately for $\mstar-$complete samples of galaxies defined by simultaneous cuts on $\bt$ and $\rhalf,$ described as follows. First, we construct subsamples of disk-dominated and bulge-dominated galaxies using cuts of $\bt<0.25$ and $\bt>0.75,$ respectively, as in \S~\ref{subsubsec:scaling_relation} above. Separately for each $\bt-$selected subsample, we calculate $\rangle\rhalf\vert\mstar\rangle,$ the median composite size of the sample as a function of total galaxy stellar mass. We split each $\bt-$selected sample in half according to whether the galaxy has a composite size $\rhalf$ that is above or below this median value. For any $\mstar-$complete threshold, this gives four subsamples of large and small, disk- and bulge-dominated galaxies. For each $\bt-$selected sample, we calculate the difference $\wproj^{\rm large}-\wproj^{\rm small},$ scaled by $\wproj^{\rm all},$ the clustering of the $\bt-$selected sample prior to splitting on $\rhalf,$ and refer to this quantity as {\em the $\rhalf$ clustering ratio}.

The y-axis of each panel in Figure \ref{fig:clustering_ratio_upshot} shows these $\rhalf$ clustering ratios, with left and right panels showing measurements for $\mstar>10^{10}M_{\odot}$ and $\mstar>10^{10.5}M_{\odot}$ thresholds in total stellar mass, and top and bottom panels showing results for disk- and bulge-dominated samples, respectively.

>>>>

In this section we present new measurements of the $\rhalf-$dependence of projected galaxy clustering, $\wproj(\rproj).$ Galaxy clustering has complex, simultaneous dependence upon $\mstar,$ $\ssfr,$ $\rhalf,$ and $\bt.$ Since the purpose of this work is to test models of composite galaxy size $\rhalf,$ we design a measurement that is specifically tailored to isolate the influence of $\rhalf$ on the two-point function, while minimizing the dependence upon the other correlated variables.

To achieve this, we compute projected clustering separately for $\mstar-$complete samples of galaxies defined by simultaneous cuts on $\bt$ and $\rhalf,$ described as follows. First, we construct subsamples of disk-dominated and bulge-dominated galaxies using cuts of $\bt<0.25$ and $\bt>0.75,$ respectively, as in \S~\ref{subsubsec:scaling_relation} above. Separately for each $\bt-$selected subsample, we calculate $\langle\rhalf\vert\mstar;\bt\rangle,$ the median composite size of the sample as a function of total galaxy stellar mass. We split each $\bt-$selected sample in half according to whether the galaxy has a composite size $\rhalf$ that is above or below this median value. For any $\mstar-$complete threshold, this gives four subsamples of large and small, disk- and bulge-dominated galaxies. For each $\bt-$selected sample, we calculate the difference $\wproj^{\rm large}-\wproj^{\rm small},$ scaled by $\wproj^{\rm all},$ the clustering of the $\bt-$selected sample prior to splitting on $\rhalf,$ and refer to this quantity as {\em the $\rhalf$ clustering ratio}.

The y-axis of each panel in Figure \ref{fig:clustering_ratio_upshot} shows these $\rhalf$ clustering ratios, with left and right panels showing measurements for $\mstar>10^{10}M_{\odot}$ and $\mstar>10^{10.5}M_{\odot}$ thresholds in total stellar mass, and top and bottom panels showing results for disk- and bulge-dominated samples, respectively. Points with error bars show SDSS measurements, solid curves show the clustering ratios of model galaxies. Before unpacking the information contained in these clustering measurements, we stress that the good agreement shown between model and data in Figure~\ref{fig:clustering_ratio_upshot} is a genuine model prediction, since we only tuned parameters to fit the one-point measurements in Fig.~\ref{fig:scatter_plot}.

The salient feature of the clustering ratio measurements is that they are negative: for either disk- or bulge-dominated samples, small galaxies cluster more strongly than large galaxies of the same stellar mass threshold, a new result. This feature also holds true for model galaxies, for which both $\rhalfbulge$ and $\rhalfdisk$ scale in power law proportion to $\rvir.$ Since halo mass $\mvir$ scales as $\rvir^{1/3},$ and clustering is a strong function of $\mvir,$ the result may seem surprising.

The resolution to this puzzle is illustrated in Figure~\ref{fig:bt_censat}, which shows the PDF of $\bt$ for halos of the same mass $\mhalo,$ defined as $\mpeak,$ the peak mass ever attained through the history of the (sub)halo, so that host halos and subhalos can be treated on equal footing. In the model, central galaxies are diskier than satellites, a feature that is inherited from the method we use to map $\bt$ to model galaxies.

In the {\tt UniverseMachine}, and in any quantitatively successful cosmological model of the galaxy formation, central galaxies are preferentially more star-forming relative to satellites of comparable stellar or halo mass. Since bluer galaxies are diskier than redder galaxies, our model centrals will be diskier than satellites. Our galaxy profile model has $\adisk=7\abulge,$ and so satellite galaxies will be smaller than centrals due to the $\bt-\ssfr$ correlation. These simultaneous features alone will naturally result in the negative sign of the clustering ratios seen in Figure~\ref{fig:clustering_ratio_upshot}.

In light of this explanation, it is natural to ask whether this simple feature {\em alone} is all that is needed for any model of $\rhalf$ to achieve this level of success at predicting galaxy clustering on small and large scales. We address this question in \S~\ref{subsec:tests2} below, finding that the answer is no: the reasonably correct magnitude and scale-dependence of the clustering ratios predicted by our model is a highly non-trivial result that places tight constraints on the post-infall physics of satellite galaxies.
